Supplement to ``Local clustering of density upper level sets''

### 2019-01-18 icml/icml_local_density_clustering_supplement.tex
completed re-organization of Theorem 2

%%% Vectors
\newcommand{\pbf}{\mathbf{p}}
\newcommand{\ebf}[1]{\mathbf{e}_{#1}}
\newcommand{\pibf}{\bm{\pi}}
\newcommand{\rhobf}{\bm{\rho}}

@@

As part of the proof of Theorem \ref{thm: inverse_mixing_time_lower_bound}, we will require an estimate of a function $g(t)$ for $t \in [x_0, 1/2]$ for some $x_0 > 0$. For $m > 0$ and $x_0 = t_0 < t_1 < \ldots < t_m = 1/2$, define the \emph{stepwise approximation to $g$} to be $\bar{g}$, given by 
\begin{equation}
\label{eqn: stepwise_approximation}
\bar{g}(t) = g(t_i), ~~ \text{ for $t \in [t_{i-1}, t_i]$ }
\end{equation}
\begin{lemma}
	\label{lem: stepwise_approximation}
	Fix
	\begin{equation*}
	g(t) = \log\left(\frac{1}{t}\right) \text{ for $x \in [x_0, 1/2]$}
	\end{equation*}
	 If for all $i$ in $1,\ldots,m$, $(t_i - t_{i - 1}) \leq x_0/2$, then $g(t) \geq \bar{g}(t)\geq g(t) / 2$.
\end{lemma}
\begin{proof}
	The upper bound $g(t) \geq \bar{g}(t)$ follows immediately from the fact that $g(t)$ is a decreasing function.

The proof of Theorem \ref{thm: inverse_mixing_time_lower_bound} also on a parameter -- which we term \emph{discrete local spread} -- to handle the mixing over very small steps. Formally, the discrete local spread $\pi_1(G)$ is given by
\begin{equation}
\label{eqn: local_spread}
\pi_1(G) := \frac{d_{\min}(G)^2}{10 \vol(V; G)} 
\end{equation}
where $d_{\min}(G) = \min_{v \in V} d(v)$ is the minimum degree in $G$. Intuitively, the discrete local spread gauges how much the walk given by $\Wbf$ has mixed after one step, starting from any node $v$. We will denote $\pi_1(G_{n,r}[\Csig[\Xbf]])$ by $\widetilde{\pi}_{1,n}$.

\begin{lemma}
	\label{lem: local_spread_lb}
	For $\Csig$ satisfying the conditions of Theorem \ref{thm: inverse_mixing_time_lower_bound}:
	\begin{equation*}
	\liminf_{n \to \infty}~ \widetilde{\pi}_{1,n} \geq \frac{\lambda_{\sigma}^2}{\Lambda_{\sigma}^2} \frac{r^d}{(2D)^d}
	\end{equation*}
\end{lemma}

\textcolor{red}{Prove Lemma \ref{lem: local_spread_lb}}.

@@

\subsection{Proof of Theorem \ref{thm: inverse_mixing_time_lower_bound}}

\textcolor{red}{Give proof structure.}

We begin by introducing some more notation. Take $G = (V,E)$ to be an undirected and unweighted graph, with associated adjacency matrix $\Abf$, random walk matrix $\Wbf$, and stationary distribution $\pibf$. We slightly overload notation and let the \emph{conductance function} be given by
\begin{equation*}
\Phi(t; G) := \min_{\substack{S \subseteq V \\ \pi(S) \leq t} } \Phi(S; G)
\end{equation*}

We will pay special attention to the conductance function evaluated over the subgraph $G_{n,r}[\Csig[\Xbf]]$. For ease of notation, we therefore introduce $\widetilde{\Phi}_{n,r}(t) = \Phi(t; G_{n,r}[\Csig[\Xbf]])$.

Lemma \ref{lem: montenegro} is from the \textcolor{red}{PhD thesis of Montenegro.} It leverages the conductance function to produce an upper bound on the \textcolor{red}{total variation distance} between the random walk 
\begin{equation*}
\rhobf^t = \frac{1}{t} \sum_{s = 1}^{t} e_v \Wbf^s
\end{equation*}
and the stationary distribution $\pibf$.

\begin{lemma}
	\label{lem: montenegro}
	Let $\Wbf$ be the random walk matrix over a graph $G = (V,E)$, with stationary distribution $\pibf$. Then, for any $v \in V$:
	\begin{equation*}
	\norm{\rho^t - \pibf}_{TV} \leq \max\left\{ \frac{1}{4}, \frac{1}{10} +  \frac{70}{t}\left(\frac{4}{\Phi^2(\pi_1(G); G)} + \int_{x = \pi_1(G)}^{1/2} \frac{4}{x \Phi^2(x; G)} dx\right) \right\}
	\end{equation*}
	where $\pi_1(G)$ is defined as in \eqref{eqn: local_spread}.
\end{lemma}

\begin{proof}
	In \textcolor{red}{the PhD thesis of Montenegro}, letting
	\begin{equation*}
	h^t(x_0) = \sup_{S: \pi(S) < x_0}  \bigl(\rhobf^t(S) - \pibf(S) \bigr)
	\end{equation*}
	the following statement is shown: for $\pi_{\min} < x_0 < 1/2$, 
	\begin{equation*}
	\norm{\rho^t - \pibf}_{TV} \leq \max\left\{ \frac{1}{4}, h^t(x_0) +  \frac{70}{t}\left(\frac{4}{\Phi^2(x_0); G)} + \int_{x = x_0}^{1/2} \frac{4}{x \Phi^2(x; G)}\right) \right\}
	\end{equation*}
	It is also observed therein that the function $h^t(x)$ is decreasing for any fixed $x$, as $t$ increases. All that remains for us to show is that $h^s(\pi_1(G)) \leq 1/10$ for some $s \leq t$. We choose $s = 1$, and note that for any set $S \subset V$ with $\pibf(S) \leq \pi_1(G)$,
	\begin{equation*}
	\rhobf^1(S) \leq \frac{\abs{S}}{d_{\min}(G)} \leq \frac{\pibf(S) \vol(V; G)}{d_{\min}(G)^2} \leq \frac{1}{10},
	\end{equation*}
	where we use the fact that $\pibf(S) \geq \frac{\abs{S} d_{\min}(G)}{\vol(V; G)}$.
\end{proof}

We turn to lower bounding $\widetilde{\Phi}_{n,r}(t)$. First, we exhibit a lower bound on a continuous space analogue, over the set $\Csig$. Let $\nu_{\Pbb}(\cdot)$ denote the weighted volume; formally, for $\Sset \subset \Rd$
\begin{equation*}
\nu_{\Pbb}(\Sset) := \int_{\Sset} f(x) dx
\end{equation*}

The $r$-ball walk over $\Csig$ is a Markov chain with transition probability for $x \in \Csig, \Sset, \Sset' \subset \Csig$ given by,
\begin{equation*}
P_{\Pbb, r}(x;\Sset) := \frac{\nu_\Pbb(\Sset \cap B(x,r))}{\nu_\Pbb(\Csig \cap B(x,r))}, ~~~ Q_{\Pbb, r}(\Sset, \Sset') := \int_{x \in \Sset} f(x) P_{\Pbb, r}(x;\Sset') dx. 
\end{equation*}
stationary distribution defined by
\begin{equation*}
\ell_{\Pbb,r}(x) := \frac{\nu_\Pbb(\Csig \cap B(x,r))}{\nu_{\Pbb}(B(x,r))}, ~~~ \pi_{\Pbb,r}(\Sset) := \frac{1}{\int_{\Csig} f(x) \ell_{\Pbb,r}(x) dx} \int_{\Sset} f(x) \ell_{\Pbb,r}(x) dx
\end{equation*}
and corresponding conductance function
\begin{equation*}
\widetilde{\Phi}_{\Pbb,r}(t) := \min_{\substack{\Sset \subset \Csig, \\ \pi_{\Pbb,r}(\Sset) \leq t} } \frac{Q_{\Pbb,r}(\Sset, \Csig \setminus \Sset)}{\pi_{\Pbb,r}(\Sset)}
\end{equation*}
Lemma \ref{lem: continuous_conductance_lb} (\textcolor{red}{a weighted analogue of Theorem 4.6 of Kannan}) provides a lower bound on $\widetilde{\Phi}_{\Pbb, r}(t)$, as well as a stepwise approximation to $\widetilde{\Phi}_{\Pbb, r}$.
\begin{lemma}
	\label{lem: continuous_conductance_lb}
	Under the conditions on $\Csig$ given by Theorem \ref{thm: inverse_mixing_time_lower_bound}, the following bounds hold: 
	\begin{itemize}
		\item 
		for $0 < t < 1/2$,
		\begin{equation*}
		\widetilde{\Phi}_{\Pbb,r}(t) > \min\left\{\frac{1}{288\sqrt{d}},\frac{r}{81 \sqrt{d}D}\log\left(\frac{\lambda_{\sigma}^2}{\Lambda_{\sigma}^2 t}\right)\right\} \cdot \frac{\lambda_{\sigma}^4}{\Lambda_{\sigma}^4}
		\end{equation*}
		\item
		Let
		\begin{equation*}
		m = \frac{2^{d+1}D^d \Lambda_{\sigma}^2}{r^d \lambda_{\sigma}^2}
		\end{equation*}
		 and $t_i = (i + 1)/m$ for $i = 0, \ldots, m - 1$. Then, for $1/m < t < 1/2$
		\begin{equation*}
		\overline{\Phi}_{\Pbb,r}(t) > \min\left\{\frac{1}{288\sqrt{d}},\frac{r}{162 \sqrt{d}D}\log\left(\frac{\Lambda_{\sigma}^2}{\lambda_{\sigma}^2 t}\right)\right\} \cdot \frac{\lambda_{\sigma}^4}{\Lambda_{\sigma}^4}
		\end{equation*}
		where $\overline{\Phi}_{\Pbb,r}(t)$ is defined as in \eqref{eqn: stepwise_approximation} with respect to $t_0, \ldots t_{m - 1}$. 
	\end{itemize}

\end{lemma}
\begin{proof}
	We begin by stating directly the result of \textcolor{red}{Kannan et al.}: for $0 < t < 1/2$, and any $S$ such that $\pi_{\nu,r}(S) = t$,
	\begin{equation}
	\label{eqn: kannan_1}
	\frac{Q_{\nu,r}(\Sset, \Csig \setminus \Sset)}{t} > \min\left\{\frac{1}{288\sqrt{d}},\frac{r}{81 \sqrt{d}D}\log\left(\frac{1}{t}\right)\right\}.
	\end{equation}
	Now, we note that
	\begin{equation*}
	\pi_{\Pbb,r}(S) \leq \pi_{\nu,r}(S) \cdot \frac{\Lambda_{\sigma}^2}{\lambda_{\sigma}^2}, ~~~ Q_{\Pbb,r}(\Sset, \Csig \setminus \Sset) \geq Q_{\nu,r}(\Sset, \Csig \setminus \Sset) \cdot \frac{\lambda_{\sigma}^2}{\Lambda_{\sigma}^2}
	\end{equation*}
	and plugging these estimates in to \eqref{eqn: kannan_1} gives the final bound on $\widetilde{\Phi}_{\Pbb,r}(t)$.  The bound on $\overline{\Phi}_{\Pbb,r}(t)$ then follows immediately from application of Lemma \ref{lem: stepwise_approximation}.
\end{proof}

>>>>

%%% Vectors
\newcommand{\pbf}{\mathbf{p}}
\newcommand{\qbf}{\mathbf{q}}
\newcommand{\ebf}[1]{\mathbf{e}_{#1}}
\newcommand{\pibf}{\bm{\pi}}
\newcommand{\rhobf}{\bm{\rho}}
\newcommand{\Deltabf}{\bm{\Delta}}
\newcommand{\deltabf}{\bm{\delta}}

%%% Random walk vectors

@@

The proof of Theorem \ref{thm: inverse_mixing_time_lower_bound} also depends on a parameter -- which we term \emph{discrete local spread} -- to handle the mixing over very small steps. Formally, the discrete local spread $\pi_1(G)$ is given by
\begin{equation}
\label{eqn: local_spread}
\pi_1(G) := \frac{d_{\min}(G)^2}{10 \vol(V; G)} 
\end{equation}
where $d_{\min}(G) = \min_{v \in V} d(v)$ is the minimum degree in $G$. Intuitively, the discrete local spread gauges how much the walk given by $\Wbf$ has mixed after one step, starting from any node $v$. We will denote $\pi_1(G_{n,r}[\Csig[\Xbf]])$ by $\widetilde{\pi}_{1,n}$.

\textcolor{red}{Prove Lemma \ref{lem: local_spread_lower_bound}}.

@@

\subsection{Mixing time on graphs}
For $N \in \mathbb{N}$ and a set $V$ of $N$ vertices, take $G = (V,E)$ to be an undirected and unweighted graph, with associated adjacency matrix $\Abf$, random walk matrix $\Wbf$, and stationary distribution $\pibf = (\pi_u)_{u \in V}$ where $\pi_v = \frac{\Dbf_{vv}}{\vol(V; G)}$. For $v \in V$, 
\begin{equation}
\label{eqn: random_walk}
q_{vu}^{(m)} = e_v\Wbf^m e_u, ~~ \qbf_{v}^{(m)} = \left(q_{vu}^{(m)}\right)_{u \in V}, ~~ \qbf_v = (\qbf^{(1)}_{v\cdot},\qbf^{(2)}_{v\cdot}, \ldots), 
\end{equation}
denote respectively the $m$-step transition probability, distribution, and sequence distributions of the random walk over $G$ originating at $v$. Letting $\qbf = (\qbf_v)_{v \in V}$, the relative pointwise mixing time is thus
\begin{equation*}
\tau_{\infty}(\qbf; G) = \min\set{m: \forall u,v \in V, \frac{\abs{q_{vu}^{(m)} - \pibf_u}}{\pibf_u} \leq 1/4} 
\end{equation*}

Two key quantities relate the mixing time to the expansion of subsets $S$ of $V$. The \emph{local spread} is defined to be
\begin{equation*}
s(G) := \frac{9D_{\min}}{10}\pi_{\min} 
\end{equation*}
for $D_{\min} := \min_{v \in V} \Dbf_{vv}$ and $\pi_{\min} := D_{\min} / \vol(V; G)$.

%\inf \set{t: \forall S \subset V ~\text{with}~ \pi(S) = t, \beta(S) \leq 1/10}
%\end{equation*}
where $\beta(S) := \inf_{v \in S} \qbf_{v}^{(1)}(S^c)$, and by convention we let $\pbf(S) = \sum_{u \in S} p_u$ for any distribution vector $\pbf = (p_u)_{u \in V}$ over $V$. We collect some necessary facts about the local spread in Lemma \ref{lem: local_spread_G}.
\begin{lemma}
	\label{lem: local_spread_G}
	\begin{itemize}
		\item If $\pibf(S) \leq s(G)$, then for every $u \in S$, $\qbf_u^{(1)}(S^c) \geq 1/10$.
		\item For any $v, u \in V$, and $m \in N$ greater than $0$, $q_{vu}^{(m)}/ \pi_{\min} \leq 1/s(G)$.
	\end{itemize}

\begin{proof}
	If $t = \pibf(S) \leq \frac{9 D_{\min}}{10} \pi_{\min}$, divide both sides by $\pi_{\min}$ to obtain
	\begin{equation*}
	\abs{S} \leq \frac{9 D_{\min}}{10}
	\end{equation*}
	which implies $\qbf_{v}^{(1)}(S^c) \geq 1/10$ for all $v \in S$. This implies the first statement.

	The second statement follows from the fact $q_{vu}^{(m)} \leq 1/D_{\min}$ for any $m$.
\end{proof}

The local spread facilitates conversion between $\tau_{\infty}(\qbf_v; G)$ and the more easily manageable \emph{total variation} mixing time, given by
\begin{equation*}
\tau_1(\rhobf; G) = \min\biggl\{m: \forall v \in V, \norm{\rhobf_v - \pibf}_{TV} \leq 1/4 \biggr\}
\end{equation*}
where 
\begin{equation}
\label{eqn: uniform_random_walk}
\rhobf_{v}^{(m)} = \frac{1}{m}\sum_{k = 1}^{m+1} \qbf_{v}^{m}, ~~ \rhobf_v = \left( \rhobf_{v}^{(1)}, \rhobf_{v}^{(2)}, \rhobf_{v}^{(3)} \ldots \right), ~~ \rhobf = \left( \rhobf_v \right)_{v \in V}
\end{equation}
and $\norm{\pbf - \pibf}_{TV} = \sum_{v \in V}\abs{p_v - \pi_v}$ is the total variation norm between distributions $\pbf$ and $\pibf$. 
\begin{lemma}
	\label{lem: tv_mixing_to_pointwise_mixing1}
	For $\qbf$ as in \eqref{eqn: random_walk} and $\rhobf$ as in \eqref{eqn: uniform_random_walk},
	\begin{equation*}
	\tau_{\infty}(\qbf; G) \leq 2752 \tau_1(\rhobf; G) \log \left(4 \max\left\{1, \frac{1}{s(G)}\right\}\right)
	\end{equation*}
\end{lemma}
\begin{proof}
	Masking dependence on the starting vertex $v$ for the moment, let
	\begin{equation*}
	\Delta_u^{(m)} = q_{vu}^{(m)} - \pi_u, ~~ \delta_u^{(m)} = \frac{\Delta_u^{(m)}}{\pi_u}
	\end{equation*}
	and $\Deltabf^{(m)} = (\Delta_u^{(m)})_{u \in V}$, $\deltabf^{(m)} = (\delta_u^{(m)})_{u \in V}$. For a vector $\Deltabf = (\Delta_u)_{u \in V}$, the $L^{p}(\pibf)$ norm is given by
	\begin{equation*}
	\norm{\Deltabf}_{L^p(\pibf)} = \left(\sum_{u \in V} \left(\Delta_u\right)^{p} \pi_u \right)^{1/p}
	\end{equation*}
	To go between the $L^{\infty}(\pibf)$ and $L^{1}(\pibf)$ norms, we have
	\begin{align*}
	\norm{\deltabf^{(2m)}}_{L^{\infty}(\pi)} & \overset{(i)}{\leq} \norm{\deltabf^{(m)}}^2_{L^{2}(\pi)} \\
	& = \norm{(\deltabf^{(m)})^2}_{L^{1}(\pi)} \\
	& \overset{(ii)}{\leq}  \norm{(\deltabf^{(m)})}_{L^{1}(\pi)} \norm{(\deltabf^{(m)})}_{L^{\infty}(\pi)}
	\end{align*}
	where $(i)$ is a result of \textcolor{red}{Benjamini and Morris} and $(ii)$ follows from Holder's inequality. Now, we upper bound the second factor on the right hand side by observing
	\begin{align*}
	\norm{(\deltabf^{(m)})}_{L^{\infty}(\pi)} & \leq \max\left\{1, \max_{u \in V} \frac{q_{vu}^{(m)}}{\pi_u} \right\} \\
	& \overset{(iii)}{\leq} \max\left\{1, \frac{1}{s(G)}\right\}
	\end{align*}
	where $(iii)$ follows from Lemma \ref{lem: local_spread_G}.

	Now, we leverage the following well-known fact (\textcolor{red}{PhD thesis of Montenegro}): for any $\epsilon > 0$, if $m \geq \tau_1(\qbf_v^{(m)}; G) \cdot \log(1/\epsilon)$ then
	\begin{equation*}
	\norm{\qbf_v^{(m)} - \pibf}_{TV} \leq \epsilon.
	\end{equation*}
	But $\norm{\qbf_v^{(m)} - \pibf}_{TV}$ is exactly $\norm{(\deltabf^{(m)})}_{L^{1}(\pi)}$. Therefore, picking 
	\begin{equation*}
	m_0 = \tau_1(\qbf_v^{(m)}; G) \cdot \log \left(4 \max\left\{1, \frac{1}{s(G)}\right\} \right)
	\end{equation*} implies $\norm{(\deltabf^{(m)})}_{L^{\infty}(\pi)} \leq 1/4$ for all $m \geq 2 m_0$.  Then, 
	\begin{equation*}
	\norm{(\deltabf^{(m)})}_{L^{\infty}(\pi)} = \sup_{u}\left\{ \frac{\abs{q_{vu}^{(m)} - \pibf_u}}{\pibf_u} \right\}.
	\end{equation*}
	and since none of the above depended on a specific choice for $v$, the supremum can be taken over all starting vertices $v$ as well. Thus $\tau_{\infty}(\qbf^{(m)}; G) \leq 2m_0$.

	Finally, it is known (\textcolor{red}{PhD thesis of Montenegro}) that
	\begin{equation*}
	\tau_{1}(\qbf^{(m)}; G) \leq 1376 \tau_{1}(\rhobf^{(m)}; G)
	\end{equation*}
	and so the desired result holds.	
\end{proof}

The second key quantity is the \emph{conductance function}
\begin{equation}
\label{eqn: conductance_function}
\Phi(t; G) := \min_{\substack{S \subseteq V, \\ \pibf(S) \leq t} } \Phi(S; G) ~~~~~~~ (\pi_{\min} \leq t < 1)
\end{equation}
where $\Phi(S; G)$ is the normalized cut of $S$ in $G$ given by \eqref{eqn: norm_cut}.

Lemma \ref{lem: expansion_to_tv_distance} leverages the conductance function and local spread to produce an upper bound on the total variation distance between $\rhobf_{v}^{(m)}$ and $\pibf.$
\begin{lemma}
	\label{lem: expansion_to_tv_distance}
	If $D_{\min} > 10$, for any $v \in V$:
	\begin{equation*}
	\norm{\rhobf_v^{(m)} - \pibf}_{TV} \leq \max\left\{ \frac{1}{4}, \frac{1}{10} +  \frac{70}{m}\left(\frac{20}{9} + \int_{t = s'(G)}^{1/2} \frac{4}{t \Phi^2(t; G)} dt\right) \right\}
	\end{equation*}
	where $s'(G) = s(G) / 9$.
\end{lemma}

To prove Lemma \ref{lem: expansion_to_tv_distance} we first introduce a generalization of $\Phi(t;G) \cdot \Phi(t;G)$ known as a blocking conductance function.
\footnote{For more details, see \textcolor{red}{PhD thesis of Montenegro}}

\begin{definition}[Blocking Conductance Function of \textcolor{red}{PhD thesis of Montenegro}]
	\label{def: blocking_conductance_function}
	For $t_0 \geq \pi_{\min}$, a function $\phi(t; G): [t_0, 1/2] \to [0,1]$ is a \emph{blocking conductance function} if for all $S \subset V$ with $\pibf(S) = t \in [t_0, 1/2]$, either of the following hold: 
	\begin{enumerate}
		\item \emph{Exterior inequality.} For all $y \in \left[\frac{1}{2}t, t\right]: \phi_{int}(S) \geq \phi(\max\{t_0,y\})$
		\item \emph{Interior inequality.} For all $y  \in \left[t, \frac{3}{2}t\right]: \phi_{ext}(S) \geq \phi(\max\{y,1 - y\})$.
	\end{enumerate}
	where $\phi_{int}$ and $\phi_{ext}$ are defined respectively as
	\begin{align*}
	\phi_{int}(S) & = \sup_{\lambda \leq \pibf(S)} \min_{\substack{B \subset S \\ \pibf(B) \leq \lambda} } \frac{\lambda \cut(S \setminus B, S^c; G)}{\vol(V; G) \left[\pibf(S) \pibf(S^c)\right]^2} \\
	\phi_{ext}(S) & = \sup_{\lambda \leq \pibf(S)} \min_{\substack{B \subset S^c \\ \pibf(B) \leq \lambda} } \frac{\lambda \cut(S \setminus B, S^c; G)}{\vol(V; G) \left[\pibf(S) \pibf(S^c)\right]^2}
	\end{align*}
\end{definition}

\begin{theorem}[\textcolor{red}{PhD thesis of Montenegro} Theorem 3.2]
	\label{thm: montenegro}
	Consider $\phi(t; G): [t_0, 1/2] \to [0,1]$ a blocking conductance function. Then, letting
	\begin{equation*}
	h^m(t_0) = \sup_{S: \pibf(S) < t_0}  \bigl(\rhobf_v^{(m)}(S) - \pibf(S) \bigr)
	\end{equation*}
	the following statement holds: if $\phi$ is a blocking conductance function,
	\begin{align*}
	\norm{\rhobf_v^{(m)} - \pibf}_{TV} & \leq \max\left\{ \frac{1}{4}, h^1(t_0) +  \frac{70}{m}\left(\frac{1}{\phi(t_0; G)} + \int_{t = t_0}^{1/2} \frac{4}{t \phi(t; G)} dt\right) \right\} \\
	\end{align*}
\end{theorem}

Note that in \textcolor{red}{PhD thesis of Montenegro} this theorem is stated with respect to $h^0$. However, in the subsequent proof it holds with respect to $h^m$, and it is observed that $h^m$ is decreasing in $m$. For our purposes it is more useful to state it with respect to $h^1$, as we have done.

\begin{proof}[Proof of Lemma \ref{lem:  expansion_to_tv_distance}]
	Consider the function $\phi_0(t, G): [s(G), 1/2] \to [0,1]$ defined by
	\begin{equation}
	\label{eqn: local_spread_blocking_conductance_function}
	\phi_0(t; G) = 
	\begin{cases}
	\frac{1}{5}, ~~~~~~~~~~~~ t = s'(G) \\
	\frac{1}{4} \Phi^2(t; G), ~ t \in \left(s'(G), 1/2\right] \\
	\end{cases}
	\end{equation}
	\begin{lemma}
		\label{lem: local_spread_blocking_conductance_function}
		If $D_{\min} > 10$, $\phi_0$ is a blocking conductance function.
	\end{lemma}
	We take Lemma \ref{lem: local_spread_blocking_conductance_function} as given, and defer the proof until after the proof of Lemma \ref{lem: expansion_to_tv_distance}.

	Lemma \ref{lem: local_spread_blocking_conductance_function} and Theorem \ref{thm: montenegro} together yield:
	\begin{equation*}
	\norm{\rhobf^t - \pibf}_{TV} \leq \max\left\{ \frac{1}{4}, h^1(s'(G)) +  \frac{70}{m}\left(5 + \int_{t = s'(G)}^{1/2} \frac{4}{t \Phi^2(t; G)}\right) \right\}
	\end{equation*}

	Then, $h^1(s'(G)) \leq 1/10$ follows exactly from the proof of Lemma \ref{lem: local_spread_G}, except now $\pibf(S) \leq s'(G)$ results in the sharper bound of $\qbf_u^{(1)}(S^c) \geq 9/10$ for every $u \in S$.
\end{proof}
\begin{proof}[Lemma \ref{lem: local_spread_blocking_conductance_function}]
	The condition $D_{\min} > 10$ ensures that $s(G) \geq \pi_{\min}$.

	It is known that $\frac{1}{4} \Phi^2(x;G)$ satisfies the exterior inequality for all $t \in (\pi_{\min}, 1/2]$.

	For $t = s'(G)$ we will instead use the interior inequality. For any $S$ such that $\pibf(S) \leq s'(G)$, the following statement holds: for every $u \in S$, $\cut(u, S^c; G) \geq 9/10 \cdot \deg(u; G)$. 
	Fixing $\lambda = \pi(S)/2$, we have
	\begin{align*}
	\phi_{int}(S) & \geq \min_{\substack{B \subset S \\ \pibf(B) \leq \lambda} } \frac{\lambda \cut(S \setminus B, S^c; G)}{\vol(V; G) \left[\lambda (1 - \lambda)\right]^2} \\
	& \geq  \min_{\substack{B \subset S \\ \pibf(B) \leq \lambda} } \frac{9 \lambda \sum_{u \in S \setminus B} \deg(u; G)}{10 \vol(V; G) \left[\lambda (1 - \lambda)\right]^2} \\
	& \geq \frac{9\lambda^2}{20[\lambda^2(1 - \lambda)^2]} \geq \frac{9}{20}. 
	\end{align*} 
\end{proof}

\subsection{Population-level conductance function.}

We will make use of the above theory with respect to the conductance function $\Phi(t; G_{n,r}[\Csig(\Xbf)])$. First, however, we introduce a population-level analogue to $\Phi(t; G_{n,r}[\Csig(\Xbf)])$ over the set $\Csig$, which we denote $\widetilde{\Phi}_{\Pbb,r}$. (In general, we will adopt the convention of using $\widetilde{f}$ to denote functionals computed with respect to $\Csig$.)

For $\Sset \subset \Rd$
\begin{equation*}
\nu_{\Pbb}(\Sset) := \int_{\Sset} f(x) dx
\end{equation*} 
is the weighted volume.

The $r$-ball walk over $\Csig$ is a Markov chain. For $x \in \Csig$ and $\Sset, \Sset' \subset \Csig$ the transition probability is given by
\begin{equation*}
\widetilde{P}_{\Pbb, r}(x;\Sset) := \frac{\nu_\Pbb(\Sset \cap B(x,r))}{\nu_\Pbb(\Csig \cap B(x,r))}, ~~~ \widetilde{Q}_{\Pbb, r}(\Sset, \Sset') := \int_{x \in \Sset} f(x) P_{\Pbb, r}(x;\Sset') dx,
\end{equation*}
stationary distribution defined by
\begin{equation*}
\ell_{\Pbb,r}(x) := \frac{\nu_\Pbb(\Csig \cap B(x,r))}{\nu_{\Pbb}(B(x,r))}, ~~~ \pi_{\Pbb,r}(\Sset) := \frac{1}{\int_{\Csig} f(x) \ell_{\Pbb,r}(x) dx} \int_{\Sset} f(x) \ell_{\Pbb,r}(x) dx
\end{equation*}
and corresponding conductance function
\begin{equation*}
\widetilde{\Phi}_{\Pbb,r}(t) := \min_{\substack{\Sset \subset \Csig, \\ \pi_{\Pbb,r}(\Sset) \leq t} } \frac{\widetilde{Q}_{\Pbb,r}(\Sset, \Csig \setminus \Sset)}{\pi_{\Pbb,r}(\Sset)}.
\end{equation*}

For $m > 0$ and $0 < t_0 < t_1 < \ldots < t_m < 1$, denote the \emph{stepwise approximation to $g$} by $\bar{g}$, defined as
\begin{equation}
\label{eqn: stepwise_approximation}
\bar{g}(t) = g(t_i), ~~ \text{ for $t \in [t_{i-1}, t_i]$ }
\end{equation}
The stepwise approximation will be important to showing the consistency results of Section \textcolor{red}{(A.7)} hold across the entire conductance function. Lemma \ref{lem: stepwise_approximation} shows that the approximation will not overly degrade our estimates of the population-level conductance function.
\begin{lemma}
	\label{lem: stepwise_approximation}
	\begin{itemize}
		\item For any function $f$ monotone decreasing in $t$ on the interval $[t_0,t_m]$, $\bar{f}(t) \leq f(t)$ for all $t \in [t_0,t_m]$.
		\item Fix
		\begin{equation*}
		g(t) = \log\left(\frac{1}{t}\right) \text{ for $x \in [t_0, 1/2]$}
		\end{equation*}
		If for all $i$ in $1,\ldots,m$, $(t_i - t_{i - 1}) \leq t_0/2$, then $\bar{g}(t)\geq g(t) / 2$.
	\end{itemize}
\end{lemma}

\begin{proof}
	The first statement is immediately obvious, and we turn to proving the second.

	The upper bound $g(t) \geq \bar{g}(t)$ follows immediately from the fact that $g(t)$ is a decreasing function along with the first statement.

\begin{theorem}[Restatement of \textcolor{red}{Kannan 2004} Theorem 4.6]
	\label{thm: kannan}
	Let $K \subset \Rd$ be a convex body of diameter $D$. Then for any $\Sset \subset K$ with $\pi_{\nu,r}(\Sset) \leq 1/2$,
	\begin{equation}
	\label{eqn: kannan}
	\frac{Q_{\nu,r}(\Sset, \Csig \setminus \Sset)}{\pi_{\nu,r}(\Sset)} > \min\left\{\frac{1}{288\sqrt{d}},\frac{r}{81 \sqrt{d}D}\log\left(1 + \frac{1}{\pi_{\nu,r}(\Sset)}\right)\right\}.
	\end{equation}
\end{theorem}

\begin{lemma}
	\label{lem: continuous_conductance_lower_bound}
	Under the conditions on $\Csig$ given by Theorem \ref{thm: inverse_mixing_time_lower_bound}, the following bounds hold: 
	\begin{itemize}
		\item 
		for $0 < t < 1/2$,
		\begin{equation*}
		\widetilde{\Phi}_{\Pbb,r}(t) > \min\left\{\frac{1}{288\sqrt{d}},\frac{r}{81 \sqrt{d}D}\log\left(1 + \frac{\lambda_{\sigma}^2}{\Lambda_{\sigma}^2 t}\right)\right\} \cdot \frac{\lambda_{\sigma}^4}{\Lambda_{\sigma}^4}
		\end{equation*}
		\item
		Let
		\begin{equation*}
		M = \frac{2^{d+1}D^d \Lambda_{\sigma}^2}{r^d \lambda_{\sigma}^2}
		\end{equation*}
		and $t_i = (i + 1)/M$ for $i = 0, \ldots, m - 1$. Then, for $1/M < t < 1/2$
		\begin{equation*}
		\overline{\Phi}_{\Pbb,r}(t) > \min\left\{\frac{1}{288\sqrt{d}},\frac{r}{162 \sqrt{d}D}\Log\left( \frac{\Lambda_{\sigma}^2}{\lambda_{\sigma}^2 t}\right)\right\} \cdot \frac{\lambda_{\sigma}^4}{\Lambda_{\sigma}^4}
		\end{equation*}
		where $\overline{\Phi}_{\Pbb,r}(t)$ is defined as in \eqref{eqn: stepwise_approximation} with respect to $t_0, \ldots t_{M - 1}$, and $\Log(A/t) = \max\{\log(1 + 2A), \log(A/t)\}$.
	\end{itemize}
\end{lemma}
Before we prove Lemma \ref{lem: continuous_conductance_lower_bound}, note that the choice of $M$ is made to ensure $t_0$ is greater than the local spread of $G_{n,r}\left[\Csig[\Xbf]\right]$, as we will see in Section \ref{sec: consistency}.
\begin{proof}[Proof of Lemma \ref{lem: continuous_conductance_lower_bound}]
	We note that
	\begin{equation*}
	\pi_{\Pbb,r}(S) \leq \pi_{\nu,r}(S) \cdot \frac{\Lambda_{\sigma}^2}{\lambda_{\sigma}^2}, ~~~ Q_{\Pbb,r}(\Sset, \Csig \setminus \Sset) \geq Q_{\nu,r}(\Sset, \Csig \setminus \Sset) \cdot \frac{\lambda_{\sigma}^2}{\Lambda_{\sigma}^2}
	\end{equation*}

	Plugging these estimates in to \eqref{eqn: kannan} gives
	\begin{equation*}
	\frac{Q_{\Pbb,r}(\Sset, \Csig \setminus \Sset)}{\pi_{\Pbb,r}(\Sset)} > \min\left\{\frac{1}{288\sqrt{d}},\frac{r}{81 \sqrt{d}D}\log\left(1 + \frac{\lambda_{\sigma}^2}{\Lambda_{\sigma}^2 \pi_{\Pbb,r}(\Sset)}\right)\right\} \cdot \frac{\lambda_{\sigma}^4}{\Lambda_{\sigma}^4}
	\end{equation*}
	and since the right hand side is decreasing in $\pi_{\Pbb,r}(\Sset)$, the desired lower bound holds on $\widetilde{\Phi}_{\Pbb,r}(t)$.  The bound on $\overline{\Phi}_{\Pbb,r}(t)$ then follows from  $\Log(A/t) \leq \log(1 + 1/t)$ for all $0 < t < 1/2$ and application of Lemma \ref{lem: stepwise_approximation}.
\end{proof}

\subsection{Consistency of local spread and conductance function.}
\label{sec: consistency}

For notational ease, we write 
\begin{equation*}
\widetilde{\Phi}_{n,r}(t) = \Phi(t; G_{n,r}[\Csig[\Xbf]]), ~~ \widetilde{s}_{n,r} = s(G_{n,r}[\Csig[\Xbf]])
\end{equation*}
